---
title: "01_Plots"
output: html_document
---

# Setup

```{r setup}
library(tidyverse)
library(terra)
library(tidyterra)
library(sf)
library(sfheaders)
library(ggspatial)
library(patchwork)
library(gdalUtilities) # ogr2ogr
library(maptiles) # for basemap
```

# Layers

## Massachusetts

```{r state_outlines}
state_outlines <- map_data("state", region = c(
    "Connecticut", "Massachusetts", "Rhode Island", "New Hampshire", "Vermont"
  )) |>
  sfheaders::sf_polygon(
    obj = _,
    x = "long",
    y = "lat",
    polygon_id = "group"
  ) |>
  `st_crs<-`(sf::st_crs("epsg:4326")) |>
  terra::vect()
ma_outline <- state_outlines |> filter(group == 3)

ma_towns <- terra::vect(file.path(
   "/vsizip",
   "data/raw",
   "MA_CENSUS2020TOWNS_SHP.zip",
   "CENSUS2020TOWNS_POLY.shp"
))
```

## ForestGEO

Data manually transcribed from https://forestgeo.si.edu/sites/north-america

```{r}
forestgeo_sites <- tribble(
  ~name, ~lat, ~lon,
  "Andrews", 46.26, -122.18,
  "Grand Teton", 43.7181, -110.7578,
  "Harvard Forest", 42.5388, -72.1755,
  "Ichauway", 21.269, -84.478,
  "Indian Cave", 40.247, -95.537,
  "Itasca State Park (ISP)", 47.2397, -95.2075,
  "Lilly Dickey Woods", 39.2359, -86.2181,
  "Michigan Big Woods", 42.4667, -84,
  "Niobara", 42.78, -100.021,
  "Ordway Swisher", 29.69101, -81.993,
  "Paint Rock", 34.772, -86.306,
  "Smithsonian Conservation Biology Institute", 38.8935, -78.1454,
  "Smithsonian Environmental Research Center", 38.8891, -76.5594,
  "Trelease Woods", 40.1297, -88.1437,
  "Tyson Research Center", 38.5178, -90.5575,
  "UNDERC (University of Notre Dame Environmental Research Center)", 46.233955, -89.535041,
  "University of California, Santa Cruz Forest Ecology Research Site (UCSC-FERP)", 37.0124, -122.075,
  "University of Maryland, Baltimore County (UMBC)", 39.2549, -76.7081,
  "Utah Forest Dynamics Plot (UFDP)", 37.6615, -112.8525,
  "Wabikon", 45.5546, -88.7945,
  "Wind River (WFDP)", 45.8197, -121.9558,
  "Yosemite", 37.7662, -119.819
) |>
  terra::vect()
```

## NEON AOP HARV

NEON flies over Harvard Forest, and much more.
The flightbox is our region of interest.

```{r neon_aop_harv}
neon_aop_harv <- terra::vect(file.path(
   "/vsizip",
   "data/raw",
   "NEON_AOP_flightBoxes_0.zip",
   "AOP_flightBoxes",
   "AOP_flightboxesAllSites.shp"
)) |>
  filter(siteID == "HARV") |>
  filter(priority == 1)
```

```{r}
ggplot() +
  geom_spatvector(
    data = state_outlines,
    color = "black",
    fill = "transparent"
  ) +
  geom_spatvector(
    data = neon_aop_harv,
    color = "red",
    fill = "red"
  ) +
  geom_spatvector(
    data = ma_towns,
    color = "blue",
    fill = "transparent"
  ) +
  coord_sf(
    crs = crs(state_outlines),
    xlim = c(ext(ma_outline)$xmin, ext(ma_outline)$xmax),
    ylim = c(ext(ma_outline)$ymin, ext(ma_outline)$ymax)
  ) +
  theme_bw() +
  ggtitle("NEON AOP HARV") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  annotation_scale(
    location = "br",
    pad_x = unit(0.45, "cm"),
    pad_y = unit(0.1, "cm")
  ) +
  annotation_north_arrow(
    location = "tr",
    width = unit(0.75, "cm"),
    pad_x = unit(-0.15, "cm"),
    pad_y = unit(0.25, "cm"),
    style = north_arrow_minimal
  )

```

Now that's settled, we can do everything with respect to the AOP.

```{r}
neon_aop_harv_xlim <- c(ext(neon_aop_harv)$xmin - 0.015, ext(neon_aop_harv)$xmax + 0.015)
neon_aop_harv_ylim <- c(ext(neon_aop_harv)$ymin - 0.015, ext(neon_aop_harv)$ymax + 0.015)

# We'll use this quite a bit too
detail_area <- terra::ext(-72.20, -72.15, 42.50, 42.55) |>
  terra::vect(crs = crs("epsg:4326"))
detail_xlim <- c(ext(detail_area)$xmin, ext(detail_area)$xmax)
detail_ylim <- c(ext(detail_area)$ymin, ext(detail_area)$ymax)

harvard_forest <- terra::vect(tibble(x=-72.1899, y=42.5319), crs=crs("epsg:4326"))

the_rest_of_the_map <- list(
  geom_spatvector(
    data = neon_aop_harv,
    aes(
      color = "NEON AOP",
      linetype = "NEON AOP"
    ),
    fill = "transparent"
  ),
  geom_spatvector(
    data = ma_towns,
    aes(
      color = "Town Borders",
      linetype = "Town Borders"
    ),
    fill = "transparent"
  ),
  geom_spatvector(
    data = detail_area,
    aes(
      color = "Detail Area",
      linetype = "Detail Area"
    ),
    fill = "transparent"
  ),
  # This one is a black background for the HF triangle;
  # it does not show in the legend.
  geom_spatvector(
    data = harvard_forest,
    color = "black",
    shape = "triangle",
    size = 2.5
  ),
  geom_spatvector(
    data = harvard_forest,
    aes(
      color = "Harvard Forest",
      linetype = "Harvard Forest"
    ),
    shape = "triangle"
  ),
  coord_sf(
    crs = crs(neon_aop_harv),
    xlim = neon_aop_harv_xlim,
    ylim = neon_aop_harv_ylim
  ),
  scale_color_manual(
    name = "Legend",
    values = c(
      "Town Borders" = "black",
      "NEON AOP" = "red",
      "Detail Area" = "blue",
      "Harvard Forest" = "gold",
      "Rope Plot" = "black"
    )
  ),
  scale_linetype_manual(
    name = "Legend",
    values = c(
      "Town Borders" = "dotted",
      "NEON AOP" = "solid",
      "Detail Area" = "dashed",
      "Harvard Forest" = "solid",
      "Rope Plot" = "solid"
    )
  ),
  theme_bw(),
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)),
  annotation_scale(
    location = "br",
    pad_x = unit(0.65, "cm"),
    pad_y = unit(0.45, "cm")
  ),
  annotation_north_arrow(
    location = "tr",
    width = unit(0.75, "cm"),
    pad_x = unit(0.00, "cm"),
    pad_y = unit(1, "cm"),
    style = north_arrow_minimal
  )
)

ggplot() +
  the_rest_of_the_map +
  ggtitle("Study Area") +
  ggplot() +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")
```

## All R9 NEON AOP

Ultimately, we're interested in all the NEON AOPs in USFS Region 9
(Northeast). Where / how many are there?

```{r}
neon_aops <- terra::vect(file.path(
  "/vsizip",
  "data/raw",
  "NEON_AOP_flightBoxes_0.zip",
  "AOP_flightBoxes",
  "AOP_flightboxesAllSites.shp"
)) |>
  filter(
#    siteType == "Core",
    sampleType == "Terrestrial",
    priority == 1
  )
# 21 of those

fsr9_regions <- tibble(region = c(
  "minnesota",
  "iowa",
  "missouri",
  "wisconsin",
  "illinois",
  "michigan",
  "indiana",
  "ohio",
  "west virginia",
  "pennsylvania",
  "new york",
  "vermont",
  "maine",
  "new hampshire",
  "massachusetts",
  "rhode island",
  "connecticut",
  "new jersey",
  "delaware",
  "maryland"
))
fsr9_regions$fsr9 = "r9"
fsr9_states_map <- map_data("state") |>
  left_join(fsr9_regions) |>
  mutate(fsr9 = coalesce(fsr9, "other"))

fsr9_states_vect <- fsr9_states_map |>
  filter(fsr9 == "r9") |>
  sf::st_as_sf(coords = c("long", "lat"), crs = "epsg:4326") |>
  group_by(group) |>
  summarize(geometry = sf::st_combine(geometry)) |>
  sf::st_cast("POLYGON") |>
  terra::vect()
```

```{r}
fsr9_xlim <- c(-100, -65)
fsr9_ylim <- c(25, 50)

fsr9_countries_map <- map_data("world", c("canada", "mexico"))

ggplot() +
  geom_polygon(
    data = fsr9_countries_map,
    mapping = aes(long, lat, group = group),
    color = "black",
    fill = "white",
    show.legend = FALSE
  ) +
  geom_polygon(
    data = fsr9_states_map,
    mapping = aes(long, lat, group = group, fill = fsr9),
    color = "black",
    show.legend = FALSE
  ) +
  geom_spatvector(
    data = forestgeo_sites |>
      terra::intersect(fsr9_states_vect),
    shape = 19,
    size = 6,
    color = "black"
  ) +
  geom_spatvector(
    data = forestgeo_sites |>
      terra::intersect(fsr9_states_vect),
    shape = 19,
    size = 5,
    color = "magenta"
  ) +
  geom_spatvector(
    data = neon_aops |>
      terra::centroids() |>
      terra::intersect(fsr9_states_vect),
    shape = 17,
    size = 4,
    color = "black"
  ) +
  geom_spatvector(
    data = neon_aops |>
      terra::centroids() |>
      terra::intersect(fsr9_states_vect),
    shape = 17,
    size = 3,
    color = "cyan"
  ) +
  coord_sf(crs = 4326, xlim = fsr9_xlim, ylim = fsr9_ylim) +
  theme_bw() +
#  theme(panel.background = element_rect(fill = "lightblue", color = NA)) +
  scale_fill_manual(values = c("r9" = "forestgreen", "other" = "white")) +
  xlab(NULL) +
  scale_x_continuous(breaks = c()) +
  ylab(NULL) +
  scale_y_continuous(breaks = c())
```

## Map Tiles

```{r}
# Download tiles
if (file.exists("data/processed/basemap.tif")) {
  basemap <- terra::rast("data/processed/basemap.tif")
} else {
  basemap_area <- terra::ext(
    neon_aop_harv_xlim["xmin"],
    neon_aop_harv_xlim["xmax"],
    neon_aop_harv_ylim["ymin"],
    neon_aop_harv_ylim["ymax"]
  ) |>
    terra::vect(crs = crs("epsg:4326"))
  
  neon_aop_web <- basemap_area |>
    terra::project(crs("epsg:3857")) |> # use webmercator
    st_as_sf()
  
  basemap <- maptiles::get_tiles(
    neon_aop_web,
    provider = "OpenStreetMap",
    zoom = 13, # 4.5km / tile
    crop = TRUE
  )
}

if (file.exists("data/processed/basemap_detail.tif")) {
  basemap_detail <- terra::rast("data/processed/basemap_detail.tif")
} else {
  detail_area_web <- detail_area |>
    terra::project(crs("epsg:3857")) |> # use webmercator
    st_as_sf()

  basemap_detail <- maptiles::get_tiles(
    detail_area_web,
    provider = "OpenStreetMap",
    zoom = 15, # 1.1km / tile
    crop =TRUE
  )
}
```

```{r}
# Plot
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  the_rest_of_the_map +
  ggtitle("Study Area") +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")
```

## Rope Plots

```{r}
rope_plots <- readxl::read_xls(
  "data/raw/Rope Plots Data Entry.xls",
  sheet = "Rope Plot Sites"
) |>
  mutate(
    lat = as.numeric(gsub("^N ?([^°]+)°.*$", "\\1", GPS)),
    lon = -as.numeric(gsub("^.*W ?([^°]+)°$", "\\1", GPS))
  ) |>
  terra::vect()
```

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatvector(
    data = rope_plots,
    aes(
      color = "Rope Plot",
      linetype = "Rope Plot"
    ),
    size = 0.75
  ) +
  the_rest_of_the_map +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  geom_spatvector(
    data = rope_plots,
    aes(
      color = "Rope Plot",
      linetype = "Rope Plot"
    ),
    size = 0.75
  ) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  plot_layout(guides = "collect")
  
```

## Aerial Imagery

There's a grid; which grid cells do we need?
```{r}
massgis_grid <- terra::vect(file.path(
   "/vsizip",
   "data/raw",
   "COQ2023INDEX_POLY.zip",
   "COQ2023INDEX_POLY.shp"
))
```

```{r}
ggplot() +
  geom_spatvector(
    data = massgis_grid |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    color = "red",
    fill = "transparent"
  ) +
  the_rest_of_the_map

```

## NLCD

The NLCD available from the MRLC viewer is much higher resolution than
the NLCD available from EarthExplorer.

* MRLC Viewer: https://www.mrlc.gov/viewer/
* Earth Explorer: https://earthexplorer.usgs.gov

MRLC generates a zipfile with the UUIDs both in the name of the
zipfile and in the names of the files contained therein. The UUID
is the identity of the job used to create the file.

Nik's Downloads:

* bddf54a4-322c-40a1-bf88-b050f871fad8: 2024
* c1a704ee-7eb3-47cc-ba90-39e87f9ee635: 1985 - 2024

```{r}
nlcd_uuid <- "c1a704ee-7eb3-47cc-ba90-39e87f9ee635"
nlcd_year <- 2024

nlcd <- terra::rast(
  file.path(
    "/vsizip",
    "data/raw",
    paste0("NLCD_", nlcd_uuid, ".zip"),
    paste0("Annual_NLCD_LndCov_", nlcd_year, "_CU_C1V1_", nlcd_uuid, ".tiff")
  )
)
```

```{r}
ggplot() +
  geom_spatraster(data = nlcd, use_coltab = TRUE) +
  the_rest_of_the_map +
  ggtitle("NLCD Land Cover Class") +
  labs(fill = "NLCD Land Cover Class")
```

## ESA Worldcover

ESA Worldcover is 10m resolution, as opposed to NLCD's 30. How's that look?

```{r}
esa_worldcover <- terra::rast(file.path(
  "/vsizip",
  "data/raw",
  "terrascope_download_20260402_090311.zip",
  "WORLDCOVER/ESA_WORLDCOVER_10M_2021_V200/MAP/ESA_WorldCover_10m_2021_v200_N42W075_Map",
  "ESA_WorldCover_10m_2021_v200_N42W075_Map.tif"
)) |>
  terra::crop(neon_aop_harv)

# from https://collections.sentinel-hub.com/worldcover/readme.html
esa_worldcover_legend <- tribble(
  ~Value, ~ColorCode, ~ClassName,
  10, "#006400", "Tree cover",
  20, "#ffbb22", "Shrubland",
  30, "#ffff4c", "Grassland",
  40, "#f096ff", "Cropland",
  50, "#fa0000", "Built up",
  60, "#b4b4b4", "Bare /sparse vegetation",
  70, "#f0f0f0", "Snow and Ice",
  80, "#0064c8", "Permanent water bodies",
  90, "#0096a0", "Herbaceous wetland",
  95, "#00cf75", "Mangroves",
  100, "#fae6a0", "Moss and lichen"
)
levels(esa_worldcover) <- esa_worldcover_legend |>
  select(Value, ClassName)
```


```{r}
ggplot() +
  geom_spatraster(data = esa_worldcover, use_coltab = TRUE) +
  the_rest_of_the_map +
  ggtitle("ESA Worldcover")
```

## NLCD ESA SxS

```{r}
ggplot() +
  geom_spatraster(data = nlcd, use_coltab = TRUE) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  ggtitle("NLCD") +
  theme(legend.position = "none") +
  ggplot() +
  geom_spatraster(data = esa_worldcover, use_coltab = TRUE) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  ggtitle("ESA") +
  theme(legend.position = "none")
```

## Ownership

I have geodatabase files for the towns that intersect the AOI; these
include parcels and Assessor information.

### File Conversion

I have Geodatabases for each town, but R crashes if I try to load
them directly. So convert them to GeoPackage.

```{r}
zipfiles <- list.files(
  file.path("data/raw"),
  ".*_parcels_gdb.zip"
)
for (zipfile in zipfiles) {
  geodatabase <- zip::zip_list(file.path("data/raw", zipfile)) |>
    filter(grepl("_sde.gdb/$", filename)) |>
    pull(filename) |>
    gsub("/$", "", x = _)
  gpkg <- gsub("_gdb.zip", ".gpkg", zipfile)
  if (!file.exists(file.path("data/processed", gpkg))) {
    gdalUtilities::ogr2ogr(
      file.path("/vsizip", "data/raw", zipfile, geodatabase),
      file.path("data/processed", gpkg),
      nlt = "CONVERT_TO_LINEAR"
    )
  }
}
```

### Merge Layers

There are two layers of interest: TaxPar, which has the parcel geometry,
and Assess, which has the Assessor record.

```{r}
load_parcels <- function(gpkg) {
  # Each GeoPackage has a bunch of layers, and they're named different
  # in each file.
  l <- terra::vector_layers(gpkg)
  # The parcels have the geometry
  parcels <- terra::vect(gpkg, layer = grepv("TaxPar$", l))
  # The assessor records have the other data we're interested in
  # A single location can have multiple assessor records, e.g.
  # a condominium will have a record for the entire association, plus
  # a separate record for each unit in the association. We take the
  # record with the largest lot size to be representative of the land
  # use.
  assess <- terra::vect(gpkg, layer = grepv("Assess$", l)) |>
    as.data.frame() |>
    # TOWN_ID is in both parcels and assess
    select(-TOWN_ID) |>
    # Some assessment records don't have a location; they make a mess,
    # so remove them.
    filter(!is.na(LOC_ID)) |>
    group_by(LOC_ID) |>
    arrange(desc(LOT_SIZE)) |>
    slice_head(n = 1) |>
    ungroup()
  # Add the assessor data to the parcels
  parcels |>
    left_join(
      assess,
      by = join_by(LOC_ID) # LOC_ID is unique within the state (in theory)
    )
}

tax_parcels <- list.files("data/processed", ".*_parcels.gpkg$", full.names = TRUE) |>
  lapply(load_parcels) |>
  terra::vect()
```

### Review

```{r}
style_residential <- tolower(c(
  "4 to 8 Family",
  "4-8 family",
  "Apartment",
  "Apartment Seniors",
  "apartments",
  "Bungalow",
  "Cape",
  "Cape Cod",
  "Colonial",
  "Condominium",
  "Contemporary",
  "contempory",
  "Conventonl",
  "Conventional",
  "cottage",
  "cottage/bungalow",
  "family conver.",
  "Family Duplex",
  "Gambrel",
  "Lodge",
  "Mobile Hm",
  "Mobile Home",
  "Modern/Contemp",
  "multi-conv",
  "Raised Ranch",
  "Ranch",
  "Salt Box",
  "Split Entry",
  "Split-Level",
  "Three Family",
  "Two Family",
  "Victorian"
))

style_commercial <- tolower(c(
  "Antique",
  "auto sales rpr",
  "Bank Branch",
  "Bar/Tavern",
  "Barbershop/Salon",
  "bed & breakfast",
  "Bowling Alley",
  "Car Dealer",
  "Cold Storage",
  "Convenience Store",
  "Discount Store",
  "Distrib Warehouse",
  "Dryclean/Laundry",
  "Fast Food Regional",
  "Gas Mart",
  "Gas Station",
  "Group Home",
  "Health Club",
  "Heavy Industrial",
  "Home for Aged",
  "Hospital",
  "hospitals-priv",
  "Kennel",
  "Laboratory",
  "Light Indust",
  "Light Industrial",
  "Lumber Storage",
  "Medical Office",
  "Mini Storage",
  "Mixed Retail/Apt",
  "Mixed Retail/Office",
  "Nursing Home",
  "Office",
  "office bldg",
  "Pharmacy",
  "R & D",
  "research/devel",
  "rest/ret/apt",
  "Restaurant",
  "retail/office",
  "Service Shop",
  "Store",
  "stores/apt com",
  "Strip Retail",
  "Supermarket",
  "Warehouse"
))

style_school <- tolower(c(
  "College",
  "Day Care Center",
  "Dormitory",
  "Elementary School",
  "High School",
  "Middle School",
  "School"
))

style_recreation <- tolower(c(
  "Camp",
  "Club/Lodge",
  "Clubs/Lodges",
  "Country Club",
  "skating arena"
))

style_municipal <- tolower(c(
  "Fire Station",
  "Library",
  "Munic/State Garage",
  "municipal garage",
  "Municipal Offices",
  "other municip",
  "Police Station",
  "post office",
  "Post Office Branch",
  "Telephone Bldg"
))

style_religious <- tolower(c(
  "churches",
  "Church/Temple",
  "convent/monastary"
))

style_farm <- tolower(c(
  "Barn",
  "Outbuildings"
))

style_vacant <- tolower(c(
  "Vacant Land"
))

parcels_with_class <- tax_parcels |>
  mutate(
    STYLE = tolower(STYLE),
    CLASS = case_when(
      STYLE %in% style_residential ~ "Residential",
      STYLE %in% style_commercial ~ "Commercial",
      STYLE %in% style_school ~ "School",
      STYLE %in% style_recreation ~ "Recreation",
      STYLE %in% style_municipal ~ "Municipal",
      STYLE %in% style_religious ~ "Religious",
      STYLE %in% style_farm ~ "Farm",
      STYLE %in% style_vacant ~ "Vacant",
      .default = STYLE
    )
  )

ggplot() +
  geom_spatvector(
    data = parcels_with_class |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    aes(fill = CLASS)
  ) +
  the_rest_of_the_map +
  scale_fill_viridis_d(
    name = "Property Type",
    breaks = c(
      "Residential",
      "Commercial",
      "School",
      "Recreation",
      "Municipal",
      "Religious",
      "Farm",
      "Vacant"
    ),
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatvector(
    data = parcels_with_class |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    aes(fill = CLASS)
  ) +
  the_rest_of_the_map +
  scale_fill_viridis_d(
    name = "Property Type",
    breaks = c(
      "Residential",
      "Commercial",
      "School",
      "Recreation",
      "Municipal",
      "Religious",
      "Farm",
      "Vacant"
    ),
    na.value = "transparent"
  ) +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  plot_layout(guides = "collect")

```

## Open Space

```{r}
# OpenSpace OWNER_TYPE decode from:
# https://www.mass.gov/info-details/massgis-data-protected-and-recreational-openspace#attributes
ma_openspace_owner_type <- "
F - Federal
S - State
C - County
M - Municipal
N - Private Nonprofit
P - Private for profit
B - Public Nonprofit
L - Land Trust
G - Conservation Organization
O - Other / None of the above (e.g.joint ownership)
X - Unknown
I - Inholding (a piece of unprotected property surrounded on all sides by a protected property or a recreational facility)
Unconfirmed:
1 - EOEEA or alternate state agencies
2 - EOEEA or non-profit
3 - EOEEA or municipality
4 - EOEEA or private landowner
" |>
  strsplit("\n") |>
  unlist() |>
  as.data.frame() |>
  set_names("text") |>
  mutate(
    OWNER_TYPE = gsub("^(.) - (.*)$", "\\1", text),
    OWNER_DESC = gsub("^(.) - (.*)$", "\\2", text)
  ) |>
  filter(!is.na(OWNER_TYPE) & OWNER_TYPE != text) |>
  select(OWNER_TYPE, OWNER_DESC) |>
  mutate(
    OWNER_NOTE = if_else(
      grepl("\\(", OWNER_DESC),
      gsub("^[^(]*\\(([^)]+)\\).*$", "\\1", OWNER_DESC),
      NA
    ),
    OWNER_DESC = if_else(
      grepl("\\(", OWNER_DESC),
      gsub("^([^(]*) ?\\(.*$", "\\1", OWNER_DESC),
      OWNER_DESC
    )
  )


# OpenSpace PUB_ACCESS decode from:
# https://www.mass.gov/info-details/massgis-data-protected-and-recreational-openspace#attributes
ma_openspace_pub_access <- "
Y - Yes (open to public)
N - No (not open to public)
L - Limited (membership only)
X - Unknown
" |>
  strsplit("\n") |>
  unlist() |>
  as.data.frame() |>
  set_names("text") |>
  mutate(
    PUB_ACCESS = gsub("^(.) - (.*)$", "\\1", text),
    PUB_ACCESS_DESC = gsub("^(.) - (.*)$", "\\2", text)
  ) |>
  filter(PUB_ACCESS_DESC != "") |>
  arrange(PUB_ACCESS) |>
  mutate(
    PUB_ACCESS_DESC2 = c("Membership Only", "Closed", "Unknown", "Open")
  )

ma_openspace <- terra::vect(file.path(
  "/vsizip",
  "data/raw",
  "MA_openspace.zip",
  "OPENSPACE_POLY.shp"
)) |>
  left_join(
    ma_openspace_owner_type, by = join_by(OWNER_TYPE)
  ) |>
  left_join(
    ma_openspace_pub_access, by = join_by(PUB_ACCESS)
  )
```

### Owner Type

```{r}
ggplot() +
  geom_spatvector(
    data = ma_openspace |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    aes(fill = OWNER_DESC)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Owner Type",
    values = c(
      "Land Trust" = "blue",
      "Municipal" = "yellow",
      "Private for profit" = "orange",
      "Private Nonprofit" = "brown",
      "State" = "green"
    ),
    breaks = c(
      "Land Trust",
      "Municipal",
      "Private for profit",
      "Private Nonprofit",
      "State"
    ),
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatvector(
    data = ma_openspace |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    aes(fill = OWNER_DESC)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Owner Type",
    values = c(
      "Land Trust" = "blue",
      "Municipal" = "yellow",
      "Private for profit" = "orange",
      "Private Nonprofit" = "brown",
      "State" = "green"
    ),
    breaks = c(
      "Land Trust",
      "Municipal",
      "Private for profit",
      "Private Nonprofit",
      "State"
    ),
    na.value = "transparent"
  ) +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  plot_layout(guides = "collect")

```

### Public Access

```{r}
ggplot() +
  geom_spatvector(
    data = ma_openspace |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    aes(fill = PUB_ACCESS_DESC2),
    alpha = 0.25
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Legend",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen",
      "Open" = "lightblue",
      "Membership Only" = "orange2",
      "Unknown" = "purple3"
    ),
    breaks = c(
      "Impervious",
      "Pervious",
      "Open",
      "Membership Only",
      "Unknown"
    ),
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatvector(
    data = ma_openspace |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    aes(fill = PUB_ACCESS_DESC2),
    alpha = 0.25
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Legend",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen",
      "Open" = "lightblue",
      "Membership Only" = "orange2",
      "Unknown" = "purple3"
    ),
    breaks = c(
      "Impervious",
      "Pervious",
      "Open",
      "Membership Only",
      "Unknown"
    ),
    na.value = "transparent"
  ) +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  plot_layout(guides = "collect")
```

## Soil

https://www.mass.gov/info-details/massgis-data-soils-ssurgo-certified-nrcs

FRMLNDCLS	Farmland Classification	Identification of map units as prime farmland, farmland of statewide importance, or farmland of unique importance.


```{r}
ma_soil_top20 <- terra::vect(
  file.path(
    "/vsizip",
    "data/raw",
    "Soils_MassGIS_GDB.zip",
    "Soils_MassGIS.gdb"
  ),
  layer = "SOILS_MUPOLYGON_TOP20"
)

ma_soil_primefarm <- terra::vect(
  file.path(
    "/vsizip",
    "data/raw",
    "Soils_MassGIS_GDB.zip",
    "Soils_MassGIS.gdb"
  ),
  layer = "SOILS_POLY_PRIMEFARMLAND"
)
```

ma_soil_primefarm contains information only about farmland;
we get "All areas are prime farmland" and "Farmland of statewide importance".
In theory there are other levels, but none appear in this AOI.

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatvector(
    data = ma_soil_primefarm |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    aes(fill = FARMLNDCL)
  ) +
  the_rest_of_the_map +
  ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatvector(
    data = ma_soil_primefarm |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    aes(fill = FARMLNDCL)
  ) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  plot_layout(guides = "collect") +
  plot_annotation(title = "Prime Farmland")

```

ma_soil_top20 has lots of soil information, including information about
farmland. In theory this should give us information about more soil
types; in practice, plotting just FRMLNDCLS (arbitrarily distinct from
FRMLNDCL in ma_soil_primefarm) gives the same answer plus a bunch of
mess from plots that aren't farmland.

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatvector(
    data = ma_soil_top20 |>
      select(FRMLNDCLS) |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv) |>
      mutate(FRMLNDCLS = if_else(FRMLNDCLS == "Not prime farmland", NA, FRMLNDCLS)),
    aes(fill = FRMLNDCLS),
    color = "transparent"
  ) +
  the_rest_of_the_map +
  scale_fill_discrete(name = "Farmland Class", na.value = "transparent")
```

Since the two give the same answer, I recommend using ma_soil_primefarm.

# Adjacency

## Adjacency Classes

We're interested in forest adjacent to impervious and pervious
surface. Collapse things down to those three, plus water. There
is no need for "Other".

```{r}
adj_classes <- tribble(
  ~ID, ~category,
  1, "Water",
  2, "Impervious",
  3, "Pervious",
  4, "Forest"
)
```

Other possible sources of candidate edges:

* MassGIS landcover - most recent is 2016
* Meta Treecover?
* Impervious product? Both MA and GHSL have them.

## NLCD to Adjacency

Reduce NLCD to three classes and NA:
- Forest (Deciduous, Evergreen, Mixed)
- Developed (Low / Medium / High Intensity)
- Open (Developed, Open Space; Barren Land; Shrub/Scrub; Grassland/Herbaceoud; Pasture/Hay; Cultivated Crops)
- NA (everything else)

```{r}
nlcd_reclassification <- tribble(
  ~is, ~becomes,
  "Open Water", "Water",
  "Developed, Open Space", "Impervious",
  "Developed, Low Intensity", "Impervious",
  "Developed, Medium Intensity", "Impervious",
  "Developed, High Intensity", "Impervious",
  "Barren Land", "Pervious",
  "Deciduous Forest", "Forest",
  "Evergreen Forest", "Forest",
  "Mixed Forest", "Forest",
  "Shrub/Scrub", "Pervious",
  "Grassland/Herbaceous", "Pervious",
  "Pasture/Hay", "Pervious",
  "Cultivated Crops", "Pervious",
  "Woody Wetlands", "Water",
  "Emergent Herbaceous Wetlands", "Water"
) |>
  left_join(levels(nlcd)[[1]], by = join_by(is == `NLCD Land Cover Class`)) |>
  left_join(adj_classes, by = join_by(becomes == category)) |>
  select(is = `Pixel Value`, becomes = `ID`)

nlcd_adj_classes <- nlcd |>
  terra::classify(nlcd_reclassification)
levels(nlcd_adj_classes) <- adj_classes
```

```{r}
nlcd_adjacency_map <- list(
  scale_fill_manual(
    name = "Land Cover Class",
    values = list(
      "Water" = "blue",
      "Impervious" = "red",
      "Pervious" = "yellow",
      "Forest" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  )
)

ggplot() +
  geom_spatraster(data = nlcd_adj_classes, use_coltab = FALSE) +
  the_rest_of_the_map +
  nlcd_adjacency_map +
  ggplot() +
  geom_spatraster(data = nlcd_adj_classes, use_coltab = FALSE) +
  the_rest_of_the_map +
  nlcd_adjacency_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  plot_layout(guides = "collect") +
  plot_annotation(title = "NLCD Collapsed Land Cover Class")
```

## ESA to Adjacency

```{r}
esa_reclassification <- tribble(
  ~is, ~becomes,
  "Tree cover", "Forest",
  "Grassland", "Pervious",
  "Cropland", "Pervious",
  "Built up", "Impervious",
  "Bare /sparse vegetation", "Pervious",
  "Permanent water bodies", "Water",
  "Herbaceous wetland", "Water"
) |>
  left_join(esa_worldcover_legend, by = join_by(is == ClassName)) |>
  left_join(adj_classes, by = join_by(becomes == category)) |>
  select(is = Value, becomes = `ID`)

esa_adj_classes <- esa_worldcover |>
  terra::classify(esa_reclassification)
levels(esa_adj_classes) <- adj_classes
```


```{r}
esa_adjacency_map <- list(
  scale_fill_manual(
    name = "Land Cover Class",
    values = list(
      "Water" = "blue",
      "Impervious" = "red",
      "Pervious" = "yellow",
      "Forest" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  )
)

ggplot() +
  geom_spatraster(data = esa_adj_classes, use_coltab = FALSE) +
  the_rest_of_the_map +
  esa_adjacency_map +
  ggplot() +
  geom_spatraster(data = esa_adj_classes, use_coltab = FALSE) +
  the_rest_of_the_map +
  esa_adjacency_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  plot_layout(guides = "collect") +
  plot_annotation(title = "ESA Collapsed Land Cover Class")
```

## NLCD ESA SxS

```{r}
ggplot() +
  geom_spatraster(data = nlcd_adj_classes, use_coltab = FALSE) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  ggtitle("Collapsed NLCD") +
  labs(fill = "Land Cover Class") +
  scale_fill_manual(
    name = "Land Cover Class",
    values = list(
      "Water" = "blue",
      "Impervious" = "red",
      "Pervious" = "yellow",
      "Forest" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  ) +
  theme(legend.position = "none") +
  ggplot() +
  geom_spatraster(data = esa_adj_classes, use_coltab = FALSE) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  ggtitle("Collapsed ESA") +
  labs(fill = "Land Cover Class") +
  scale_fill_manual(
    name = "Land Cover Class",
    values = list(
      "Water" = "blue",
      "Impervious" = "red",
      "Pervious" = "yellow",
      "Forest" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  ) +
  theme(legend.position = "none")
```

## Edges

Can use terra::adjacent to find just N/S adjacency.

looking at the revised landcover map, its one giant patch of forest, so thats not useful

instead, look for e.g. nonforest south and forest north
or, two passes: find southern edge, then find north of southern edge - will be much smaller

South-facing edges adjacent to impervious

```{r}
# Pervious Edges
pvn <- 2^0
pvs <- 2^1
pve <- 2^2
pvw <- 2^3

# Impervious Edges
ipw <- 2^4
ipe <- 2^5
ips <- 2^6
ipn <- 2^7

edge_detect <- function(x, ...) {
  lc_water <- 1
  lc_impervious <- 2
  lc_pervious <- 3
  lc_forest <- 4
  
  lc_north <- x[[2]]
  lc_west <- x[[4]]
  lc_center <- x[[5]]
  lc_east <- x[[6]]
  lc_south <- x[[8]]

  if (is.na(lc_center) | lc_center != lc_forest) {
    return(NA)
  }
  
  edges = 0
  if (!is.na(lc_north) & !is.na(lc_south)) {
    if (lc_north == lc_forest) {
      if (lc_south == lc_impervious) {
        edges <- edges + ips
      } else if (lc_south == lc_pervious) {
        edges <- edges + pvs
    }
    } else if (lc_south == lc_forest) {
      if (lc_north == lc_impervious) {
        edges <- edges + ipn
      }
      else if (lc_north == lc_pervious) {
        edges <- edges + pvn
      }
    }
  }

  if (!is.na(lc_east) & !is.na(lc_west)) {
    if (lc_east == lc_forest) {
      if (lc_west == lc_impervious) {
        edges <- edges + ipw
      }
      else if (lc_west == lc_pervious) {
        edges <- edges + pvw
      }
    } else if (lc_west == lc_forest) {
      if (lc_east == lc_impervious) {
        edges <- edges + ipe
      }
      else if (lc_east == lc_pervious) {
        edges <- edges + pve
      }
    }
  }

  if (edges == 0) {
    return(NA)
  } else {
    return(edges)
  }
}
```

```{r}
edge_matrix <- tribble(
  ~Value, ~Edge,
  pvn, "N Perv",
  pvs, "S Perv",
  pve, "E Perv",
  pvw, "W Perv",
  
  pvn + pve, "N Perv / E Perv",
  pvn + pvw, "N Perv / W Perv",
  pvs + pve, "S Perv / E Perv",
  pvs + pvw, "S Perv / W Perv",

  pvn + ipe, "N Perv / E Imp",
  pvn + ipw, "N Perv / W Imp",
  pvs + ipe, "S Perv / E Imp",
  pvs + ipw, "S Perv / W Imp",

  ipn + pve, "N Imp / E Perv",
  ipn + pvw, "N Imp / W Perv",
  ips + pve, "S Imp / E Perv",
  ips + pvw, "S Imp / W Perv",
  
  ipn + ipe, "N Imp / E Imp",
  ipn + ipw, "N Imp / W Imp",
  ips + ipe, "S Imp / E Imp",
  ips + ipw, "S Imp / W Imp",

  ipn, "N Imp",
  ips, "S Imp",
  ipe, "E Imp",
  ipw, "W Imp"
)

r <- rast(ncols = 17, nrow = 5, ext(1, 17, 1, 5))
values(r) <- c(
  rep(4, 17), # Forest
  rep(4, 17), # Forest
  c(4, 4, 1, 4, 4, 2, 4, 4, 3, 4, 4, 4, 4, 4, NA, 4, 4),
  rep(4, 17), # Forest
  rep(4, 17) # Forest
)
edges <- terra::focal(r, w = 3, fun = edge_detect, na.policy = "omit")
levels(edges) <- edge_matrix
ggplot() +
  geom_spatraster(data = edges) +
  coord_sf()
```

### NLCD Edge Type & Direction

```{r}
nlcd_edges <- nlcd_adj_classes |>
  terra::focal(w = 3, fun = edge_detect, na.policy = "omit")
levels(nlcd_edges) <- edge_matrix
```

```{r}
ggplot() +
  geom_spatraster(data = nlcd_edges) +
  the_rest_of_the_map +
  scale_fill_viridis_d(
    name = "Edge Direction & Type",
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatraster(data = nlcd_edges) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  scale_fill_viridis_d(
    name = "Edge Direction & Type",
    na.value = "transparent"
  ) +
  theme(legend.position = "none") +
  plot_layout(guides = "collect")

```


### NLCD Edge Type

```{r}
nlcd_edges_pi <- nlcd_edges |>
  mutate(Edge = if_else(
    bitwAnd(Edge, ipn + ips + ipe + ipw) != 0, 1, if_else(
      bitwAnd(Edge, pvn + pvs + pve + pvw) != 0, 2,
      NA
    ),
  ))
levels(nlcd_edges_pi) <- tibble(Value = c(1, 2), Adjacent = c("Impervious", "Pervious"))
```

```{r}
ggplot() +
  geom_spatraster(data = nlcd_edges_pi, use_coltab = FALSE) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Edge Type",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = c("Impervious", "Pervious"),
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatraster(data = nlcd_edges_pi) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  scale_fill_manual(
    name = "Edge Type",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = c("Impervious", "Pervious"),
    na.value = "transparent"
  ) +
  plot_layout(guides = "collect")

```

### NLCD vs Rope


```{r}
ggplot() +
  geom_spatraster(data = nlcd_edges_pi, use_coltab = FALSE) +
  geom_spatvector(
    data = rope_plots,
    aes(
      color = "Rope Plot",
      linetype = "Rope Plot"
    ),
    size = 0.75
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Edge Type",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = c("Impervious", "Pervious"),
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatraster(data = nlcd_edges_pi) +
  geom_spatvector(
    data = rope_plots,
    aes(
      color = "Rope Plot",
      linetype = "Rope Plot"
    ),
    size = 0.75
  ) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  scale_fill_manual(
    name = "Edge Type",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = c("Impervious", "Pervious"),
    na.value = "transparent"
  ) +
  plot_layout(guides = "collect")

```

### NLCD N/S

This doesn't work; it's of the opinion that there are no N/S Pervious
edges, which is wrong. It's also of the opinion that the long, vertical
edge along the west side of the road in the center of the detail area
is a N/S stack of N/S edges, which is impossible.

```{r}
nlcd_edges_ns <- nlcd_edges |>
  mutate(Edge = if_else(
    bitwAnd(Edge, ipn + ips) != 0, 1, if_else(
      bitwAnd(Edge, pvn + pvs) != 0, 2,
      NA
    ),
  ))
levels(nlcd_edges_ns) <- tibble(Value = c(1, 2), Adjacent = c("Impervious", "Pervious"))
```

```{r}
ggplot() +
  geom_spatraster(data = nlcd_edges_ns) +
  the_rest_of_the_map +
  scale_fill_viridis_d(
    name = "North/South Edge Type",
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatraster(data = nlcd_edges_ns) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  scale_fill_viridis_d(
    name = "North/South Edge Type",
    na.value = "transparent"
  ) +
  theme(legend.position = "none") +
  plot_layout(guides = "collect")


```


### ESA Edge Type & Direction

```{r}
esa_edges <- esa_adj_classes |>
  terra::focal(w = 3, fun = edge_detect, na.policy = "omit")
levels(esa_edges) <- edge_matrix
```

```{r}
ggplot() +
  geom_spatraster(data = esa_edges) +
  the_rest_of_the_map +
  scale_fill_viridis_d(
    name = "Edge Direction & Type",
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatraster(data = esa_edges) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  scale_fill_viridis_d(
    name = "Edge Direction & Type",
    na.value = "transparent"
  ) +
  theme(legend.position = "none") +
  plot_layout(guides = "collect")

```

### ESA Edge Type

```{r}
esa_edges_pi <- esa_edges |>
  mutate(Edge = if_else(
    bitwAnd(Edge, ipn + ips + ipe + ipw) != 0, 1, if_else(
      bitwAnd(Edge, pvn + pvs + pve + pvw) != 0, 2,
      NA
    ),
  ))
levels(esa_edges_pi) <- tibble(Value = c(1, 2), Adjacent = c("Impervious", "Pervious"))
```

```{r}
ggplot() +
  geom_spatraster(data = esa_edges_pi, use_coltab = FALSE) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Edge Type",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = c("Impervious", "Pervious"),
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatraster(data = esa_edges_pi) +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  scale_fill_manual(
    name = "Edge Type",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = c("Impervious", "Pervious"),
    na.value = "transparent"
  ) +
  plot_layout(guides = "collect")

```

# Overlays

## Edge Access

Putting edges and access together

```{r}
accessible_edges <- nlcd_edges_pi |>
  terra::intersect(
    ma_openspace |>
      filter(PUB_ACCESS == "Y" | PUB_ACCESS == "M") |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv) |>
      terra::project(crs(nlcd_edges_pi))
  )
```

```{r}
edge_access_map <- list(
  geom_spatvector(
    data = ma_openspace |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv) |>
      filter(
        PUB_ACCESS_DESC2 != "Closed"
      ),
    mapping = aes(
      fill = PUB_ACCESS_DESC2
    ),
    alpha = 0.25
  ),
  geom_spatraster(
    data = nlcd_edges_pi,
    use_coltab = FALSE
  ),
  scale_fill_manual(
    name = "Legend",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen",
      "Open" = "lightblue",
      "Membership Only" = "orange2",
      "Unknown" = "purple3"
    ),
    breaks = c(
      "Impervious",
      "Pervious",
      "Open",
      "Membership Only",
      "Unknown"
    ),
    na.value = "transparent"
  )
)
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  edge_access_map +
  the_rest_of_the_map +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  edge_access_map +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  plot_layout(guides = "collect")
```


## Soil Type


```{r}
edge_soil_map <- list(
  geom_spatvector(
    data = ma_soil_primefarm |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv) |>
      mutate(FARMLNDCL = recode_values(
        FARMLNDCL,
        "All areas are prime farmland" ~ "Prime farmland",
        "Farmland of statewide importance" ~ "Other farmland"
      )),
    aes(fill = FARMLNDCL)
  ),
  geom_spatraster(
    data = nlcd_edges_pi,
    use_coltab = FALSE
  ),
  scale_fill_manual(
    name = "Legend",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen",
      "Prime farmland" = "lightblue",
      "Other farmland" = "orange2"
    ),
    breaks = c(
      "Impervious",
      "Pervious",
      "Prime farmland",
      "Other farmland"
    ),
    na.value = "transparent"
  )
)
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  edge_soil_map +
  the_rest_of_the_map +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  edge_soil_map +
  the_rest_of_the_map +
  coord_sf(crs = crs("epsg:4326"), xlim = detail_xlim, ylim = detail_ylim) +
  plot_layout(guides = "collect")
```

