---
title: "01_Plots"
output: html_document
---

# Setup

```{r setup}
library(tidyverse)
library(terra)
library(tidyterra)
library(sf)
library(sfheaders)
library(ggspatial)
library(gdalUtilities) # ogr2ogr
library(foreign) # read.dbf
```

# Layers

## Massachusetts

```{r state_outlines}
state_outlines <- map_data("state", region = c(
    "Connecticut", "Massachusetts", "Rhode Island", "New Hampshire", "Vermont"
  )) |>
  sfheaders::sf_polygon(
    obj = _,
    x = "long",
    y = "lat",
    polygon_id = "group"
  ) |>
  `st_crs<-`(sf::st_crs("epsg:4326")) |>
  terra::vect()
ma_outline <- state_outlines |> filter(group == 3)

ma_towns <- terra::vect(
  "/vsizip/data/raw/MA_CENSUS2020TOWNS_SHP.zip/CENSUS2020TOWNS_POLY.shp"
)
```

## NEON AOP HARV

NEON flies over Harvard Forest, and much more.
The flightbox is our region of interest.

```{r neon_aop_harv}
neon_aop_harv <- terra::vect(
  "/vsizip/data/raw/NEON_AOP_flightBoxes_0.zip/AOP_flightBoxes/AOP_flightboxesAllSites.shp"
) |>
  filter(siteID == "HARV") |>
  filter(priority == 1)
```

```{r}
ggplot() +
  geom_spatvector(
    data = state_outlines,
    color = "black",
    fill = "transparent"
  ) +
  geom_spatvector(
    data = neon_aop_harv,
    color = "red",
    fill = "red"
  ) +
  geom_spatvector(
    data = ma_towns,
    color = "blue",
    fill = "transparent"
  ) +
  coord_sf(
    crs = crs(state_outlines),
    xlim = c(ext(ma_outline)$xmin, ext(ma_outline)$xmax),
    ylim = c(ext(ma_outline)$ymin, ext(ma_outline)$ymax)
  ) +
  theme_bw() +
  ggtitle("NEON AOP HARV") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  annotation_scale(
    location = "br",
    pad_x = unit(0.45, "cm"),
    pad_y = unit(0.1, "cm")
  ) +
  annotation_north_arrow(
    location = "tr",
    width = unit(0.75, "cm"),
    pad_x = unit(-0.15, "cm"),
    pad_y = unit(0.25, "cm"),
    style = north_arrow_minimal
  )

```

Now that's settled, we can do everything with respect to the AOP.

```{r}
neon_aop_harv_xlim = c(ext(neon_aop_harv)$xmin - 0.015, ext(neon_aop_harv)$xmax + 0.015)
neon_aop_harv_ylim = c(ext(neon_aop_harv)$ymin - 0.015, ext(neon_aop_harv)$ymax + 0.015)
the_rest_of_the_map <- list(
  geom_spatvector(
    data = neon_aop_harv,
    aes(
      color = "NEON AOP",
      linetype = "NEON AOP"
    ),
    fill = "transparent"
  ),
  geom_spatvector(
    data = ma_towns,
    aes(
      color = "Town Borders",
      linetype = "Town Borders"
    ),
    fill = "transparent"
  ),
  coord_sf(
    crs = crs(neon_aop_harv),
    xlim = neon_aop_harv_xlim,
    ylim = neon_aop_harv_ylim
  ),
  scale_color_manual(
    name = "Legend",
    values = c(
      "Town Borders" = "black",
      "NEON AOP" = "red"
    )
  ),
  scale_linetype_manual(
    name = "Legend",
    values = c(
      "Town Borders" = "dotted",
      "NEON AOP" = "solid"
    )
  ),
  theme_bw(),
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)),
  annotation_scale(
    location = "br",
    pad_x = unit(0.65, "cm"),
    pad_y = unit(0.45, "cm")
  ),
  annotation_north_arrow(
    location = "tr",
    width = unit(0.75, "cm"),
    pad_x = unit(0.00, "cm"),
    pad_y = unit(1, "cm"),
    style = north_arrow_minimal
  )
)
ggplot() +
  the_rest_of_the_map +
  ggtitle("Study Area")
```

## NLCD

```{r}
nlcd <- terra::rast(
  file.path(
    "/vsizip",
    "data/raw",
    "NLCD_bddf54a4-322c-40a1-bf88-b050f871fad8.zip",
    "Annual_NLCD_LndCov_2024_CU_C1V1_bddf54a4-322c-40a1-bf88-b050f871fad8.tiff"
  )
)
```

```{r}
ggplot() +
  geom_spatraster(data = nlcd, use_coltab = TRUE) +
  the_rest_of_the_map +
  ggtitle("Land Cover Class") +
  labs(fill = "Land Cover Class")
```

## Ownership

I have geodatabase files for the towns that intersect the AOI; these
include parcels and Assessor information.

I have Geodatabases for each town, but R crashes if I try to load
them directly. So convert them to Shapefiles.

```{r}
zipfiles <- list.files(
  file.path("data/raw"),
  ".*_parcels_gdb.zip"
)
for (zipfile in zipfiles) {
  geodatabase <- zip::zip_list(file.path("data/raw", zipfile)) |>
    filter(grepl("_sde.gdb/$", filename)) |>
    pull(filename) |>
    gsub("/$", "", x = _)
  shapefile <- gsub("_gdb.zip", ".shp", zipfile)
  gdalUtilities::ogr2ogr(
    file.path("/vsizip", "data/raw", zipfile, geodatabase),
    file.path("data/processed", shapefile),
    append = TRUE
  )
}
```


```{r}
load_parcels <- function(shapedir) {
  municipality <- gsub("_parcels.shp", "", shapedir)
  # Layers: TaxPar, Assess, _LUT, OthLeg, UC_LUT, Misc
  # We need TaxPar, which has the actual parcels, and
  # Assess, which is no longer a layer, but has USE_CODE
  assessment <- file.path(
    "data/processed",
    shapedir,
    paste0(municipality, "Assess.dbf")
  ) |>
    foreign::read.dbf()
  shapefile <- paste0(municipality, "TaxPar.shp")
  terra::vect(
    file.path("data/processed", shapedir, shapefile)
  ) |>
    left_join(
      assessment |>
        select(-c(TOWN_ID, PROP_ID)),
      by = join_by(LOC_ID)
    )
}

tax_parcels <- list.files("data/processed", ".*_parcels.shp$") |>
  lapply(load_parcels) |>
  terra::vect()
```

```{r}
style_residential <- c(
  "4 to 8 Family",
  "Apartment",
  "Apartment Seniors",
  "Bungalow",
  "Cape Cod",
  "Colonial",
  "Condominium",
  "Conventional",
  "Family Duplex",
  "Mobile Home",
  "Modern/Contemp",
  "Raised Ranch",
  "Ranch",
  "Split-Level",
  "Three Family",
  "Two Family",
  "Victorian"
)

style_commercial <- c(
  "Antique",
  "Bank Branch",
  "Bar/Tavern",
  "Barbershop/Salon",
  "Bowling Alley",
  "Car Dealer",
  "Cold Storage",
  "Convenience Store",
  "Discount Store",
  "Distrib Warehouse",
  "Dryclean/Laundry",
  "Fast Food Regional",
  "Gas Mart",
  "Gas Station",
  "Group Home",
  "Health Club",
  "Heavy Industrial",
  "Home for Aged",
  "Hospital",
  "Kennel",
  "Laboratory",
  "Light Industrial",
  "Lumber Storage",
  "Medical Office",
  "Mini Storage",
  "Mixed Retail/Apt",
  "Mixed Retail/Office",
  "Nursing Home",
  "Office",
  "Pharmacy",
  "R & D",
  "Restaurant",
  "Service Shop",
  "Store",
  "Strip Retail",
  "Supermarket",
  "Warehouse"
)

style_school <- c(
  "College",
  "Day Care Center",
  "Elementary School",
  "High School",
  "Middle School"
)

style_recreation <- c(
  "Camp",
  "Club/Lodge",
  "Country Club"
)

style_municipal <- c(
  "Fire Station",
  "Library",
  "Munic/State Garage",
  "Municipal Offices",
  "Police Station",
  "Post Office Branch",
  "Telephone Bldg"
)

#  "Church/Temple"     
#"Outbuildings"     
#"Parking Garage"     
#"Vacant Land"  

x <- tax_parcels |>
  mutate(
    CLASS = case_when(
      STYLE %in% style_residential ~ "Residential",
      STYLE %in% style_commercial ~ "Commercial",
      STYLE %in% style_school ~ "School",
      STYLE %in% style_recreation ~ "Recreation",
      STYLE %in% style_municipal ~ "Municipal",
      .default = STYLE
    )
  )

ggplot() +
  geom_spatvector(
    data = x |>
      terra::project(neon_aop_harv) |>
      terra::intersect(neon_aop_harv),
    aes(fill = CLASS)
  ) +
  the_rest_of_the_map
```

