---
title: "02_Viz"
output: html_document
---

# Setup

```{r setup}
library(targets)
library(tarchetypes)
library(tidyverse)
library(terra)
library(tidyterra)
library(ggspatial)
library(patchwork)
tar_source()
tar_objects_defined_in_rmd("00_DataProcessing.Rmd") |>
  tar_load()
```

# Base Map

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  the_rest_of_the_map +
  ggtitle("Study Area") +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  the_rest_of_the_map +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  ) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")
```

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatvector(data = qsq) +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  )

```

# Simplified Landcover


```{r}
ggplot() +
  #geom_spatraster_rgb(data = basemap) +
  geom_spatraster(
    data = lc4 |>
      `levels<-`(lc4_classes)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Land Cover Class",
    values = list(
      "Water" = "blue",
      "Impervious" = "red",
      "Pervious" = "yellow",
      "Forest" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  ) +
  ggtitle("Study Area") +
  ggplot() +
  #geom_spatraster_rgb(data = basemap_detail) +
  geom_spatraster(
    data = lc4 |>
      `levels<-`(lc4_classes)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Land Cover Class",
    values = list(
      "Water" = "blue",
      "Impervious" = "red",
      "Pervious" = "yellow",
      "Forest" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  ) +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  ) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")
```

# Edges

## Find the Edges

```{r}
edges <- lc4 |>
  terra::focal(
    w = 3, # 3x3 window
    fun = "min", # 4 if all forest
    na.rm = TRUE
  ) |>
  filter(focal_min < 4) |>
  terra::mask(lc4 |> filter(category == 4))
```


```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(
    data = edges |>
      `levels<-`(lc4_classes)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Land Cover Class",
    values = list(
      "Water" = "blue",
      "Impervious" = "red",
      "Pervious" = "yellow",
      "Forest" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  ) +
  ggtitle("Study Area") +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  geom_spatraster(
    data = edges |>
      `levels<-`(lc4_classes)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Land Cover Class",
    values = list(
      "Water" = "blue",
      "Impervious" = "red",
      "Pervious" = "yellow",
      "Forest" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  ) +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  ) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")
```

## Distance to Edge

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(
    data = edge_distance |>
      `levels<-`(edge_classes)
  ) +
  the_rest_of_the_map +
  ggtitle("Study Area") +
  scale_fill_manual(
    values = list(
      "Edge" = "orange",
      "Transition" = "green",
      "Interior" = "forestgreen"
    ),
    breaks = list(
      "Non-Forest",
      "Edge",
      "Transition",
      "Interior"
    ),
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  geom_spatraster(
    data = edge_distance |>
      `levels<-`(edge_classes)
  ) +
  the_rest_of_the_map +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  ) +
  ggtitle("Detail Area") +
  scale_fill_manual(
    values = list(
      "Edge" = "orange",
      "Transition" = "green",
      "Interior" = "forestgreen"
    ),
    breaks = list(
      "Non-Forest",
      "Edge",
      "Transition",
      "Interior"
    ),
    na.value = "transparent"
  ) +
  plot_layout(guides = "collect")
```

## Streams & Wetlands

We don't want edges that are close to wetlands.

MassGIS wetlands does not include streams.

MassGIS rivers and streams only has 6 streams in the NEON HARV AOP,
and they're all covered by other wetlands products.

See: https://www.mass.gov/info-details/massgis-data-massdep-hydrography-125000#attributes

```{asis}
ma_streams <- terra::vect(file.path(
  "/vsizip",
  "data/raw",
  "hydro25k.zip",
  "HYDRO25K_POLY.shp"
))
```

It seems that, if I want streams, I may need to turn to OpenTopoMap.

Just use wetlands for now; hopefully streams are small enough to not matter.

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(
    data = wetland_adj |>
      `levels<-`(wetland_adj_classes)
  ) +
  the_rest_of_the_map +
  ggtitle("Study Area") +
  scale_fill_manual(
    values = c(
      "Wet" = "red",
      "Dry" = "forestgreen"
    ),
    breaks = c(
      "Wet",
      "Dry"
    ),
    na.value = "transparent"
  ) +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  geom_spatraster(
    data = wetland_adj |>
      `levels<-`(wetland_adj_classes)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    values = c(
      "Wet" = "red",
      "Dry" = "forestgreen"
    ),
    breaks = c(
      "Wet",
      "Dry"
    ),
    na.value = "transparent"
  ) +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  ) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")
  
```

## Interior Adjacent

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(
    data = interior_adj |>
      `levels<-`(interior_adj_classes)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    values = list(
      "Isolated" = "red",
      "Interior Adjacent" = "forestgreen"
    ),
    breaks = list(
      "Interior Adjacent",
      "Isolated"
    ),
    na.value = "transparent"
  ) +
  ggtitle("Study Area") +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  geom_spatraster(
    data = interior_adj |>
      `levels<-`(interior_adj_classes)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    values = list(
      "Isolated" = "red",
      "Interior Adjacent" = "forestgreen"
    ),
    breaks = list(
      "Interior Adjacent",
      "Isolated"
    ),
    na.value = "transparent"
  ) +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  ) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")
```


## Adjacent Surface

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(
    data = impervious_adj |>
      `levels<-`(lc4_classes)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = list(
      "Impervious",
      "Pervious"
    ),
    na.value = "transparent"
  ) +
  ggtitle("Study Area") +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  geom_spatraster(
    data = impervious_adj |>
      `levels<-`(lc4_classes)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = list(
      "Impervious",
      "Pervious"
    ),
    na.value = "transparent"
  ) +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  ) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")
```

## Aspect

NOTE: Aspect doesn't work. It looks like terra::adjacent() is giving
cells out of order or something; the results just don't make sense.

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(
    data = edge_aspect |>
      mutate(category = case_when(
        category == 1 ~ "north",
        category == -1 ~ "south",
        category == 0 ~ "other"
      ))
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    values = c(
      "north" = "red",
      "south" = "forestgreen",
      "other" = "yellow"
    ),
    breaks = c(
      "north", "south", "other"
    ),
    na.value = "transparent"
  ) +
  ggtitle("Study Area") +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  geom_spatraster(
    data = edge_aspect |>
      mutate(category = case_when(
        category == 1 ~ "north",
        category == -1 ~ "south",
        category == 0 ~ "other"
      ))
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    values = c(
      "north" = "red",
      "south" = "forestgreen",
      "other" = "yellow"
    ),
    breaks = c(
      "north", "south", "other"
    ),
    na.value = "transparent"
  ) +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  ) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")

```

## Final Edges

We want:

- N/S facing edges
- Not near water
- Connected to interior forest
- With adjacent land cover

```{r}
final_edges <- c(
    impervious_adj |> select(adjacent = category),
    edge_aspect |> select(aspect = category)
  ) |>
  terra::mask(interior_adj == 1, maskvalues = 0) |>
  terra::mask(wetland_adj == 1, maskvalues = 0) |>
  terra::mask(edge_distance |> filter(category == 1)) |>
  `levels<-`(lc4_classes)
```

```{r}
ggplot() +
  geom_spatraster_rgb(data = basemap) +
  geom_spatraster(
    data = final_edges,
    aes(fill = category)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Adjacent Landcover",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  ) +
  ggtitle("Study Area") +
  ggplot() +
  geom_spatraster_rgb(data = basemap_detail) +
  geom_spatraster(
    data = final_edges,
    aes(fill = category)
  ) +
  the_rest_of_the_map +
  scale_fill_manual(
    name = "Adjacent Landcover",
    values = list(
      "Impervious" = "red",
      "Pervious" = "forestgreen"
    ),
    breaks = c(
      "Water",
      "Impervious",
      "Pervious",
      "Forest"
    ),
    na.value = "transparent"
  ) +
  coord_sf(
    crs = crs("epsg:4326"),
    xlim = c(-72.20, -72.15),
    ylim = c(42.50, 42.55)
  ) +
  ggtitle("Detail Area") +
  plot_layout(guides = "collect")
```
