---
title: "04_LaserPlots"
output: html_document
---

# Setup

```{r setup}
library(tidyverse)
library(terra)
library(tidyterra)
library(allodb)

for (script in list.files(path = "R", pattern = "*.[Rr]", full.names = TRUE)) {
  source(script)
}
```

# Datums

```{r}
species <- readxl::read_xlsx("data/raw/Edge Plot Data Sheet.xlsx", sheet = "Tree Species") |>
  mutate(genspe= paste0(toupper(substr(GENUS, 1, 3)), toupper(substr(SPECIES, 1, 3))))
trees <- readxl::read_xlsx("data/raw/Edge Plot Data Sheet.xlsx", sheet = "Tree Data") |>
  left_join(species, by = join_by(gesp))
plots <- readxl::read_xlsx("data/raw/Edge Plot Data Sheet.xlsx", sheet = "Plot Data") |>
  mutate(
    lat = dms_to_dd(lat),
    lon = dms_to_dd(lon)
  )
harvard_forest <- terra::vect(file.path(
  "/vsizip",
  "data/processed",
  "harvard_forest.shp.zip",
  "harvard_forest.shp.shp"
)) |>
  as.data.frame(geom = "XY") |>
  select(lon = x, lat = y) |>
  as.list()
```

## Datums for Garmin

Turn the plots into a GPX file for Garmin GPS.
For GPX, the layer name matters: point features must be written to the waypoints layer. GPX uses WGS 84 coordinates, so EPSG:4326 is correct. The GDAL GPX driver supports point features as waypoints, lines as routes, and multilines as tracks.

Preparing the waypoint attributes

The most useful standard GPX fields are:

name: waypoint name shown on the Garmin
desc: description
sym: Garmin/GPX symbol name
ele: elevation in metres
time: timestamp

```{r, eval=FALSE}
waypoints <- plots |>
  select(
    name = plot_id,
    desc = edge,
    time = date,
    lon,
    lat
  ) |>
  terra::vect(crs = crs("epsg:4326"))
waypoints |>
  terra::writeVector(
    "EstimatedPlots.gpx",
    filetype = "GPX",
    layer = "waypoints"
  )
```

# Data Validation

```{r}
# All plots must have trees
stopifnot(
  plots |>
    distinct(plot_id) |>
    anti_join(trees |> distinct(plot_id)) |>
    summarize(n = n()) |>
    pull(n) == 0
)

# All plots must have metadata
stopifnot(
  trees |>
    distinct(plot_id) |>
    anti_join(plots |> distinct(plot_id)) |>
    summarize(n = n()) |>
    pull(n) == 0
)

# All plots must have all 3 subplots represented
# not technically true, but we want to carefully examine cases when it's
# false
stopifnot(
  trees |>
    distinct(plot_id, subplot) |>
    group_by(plot_id) |>
    summarize(n = n()) |>
    ungroup() |>
    filter(n < 3) |>
    summarize(n = n()) |>
    pull(n) == 0
)

# All species must have metadata
stopifnot(
  trees |>
    filter(gesp != "SNAG") |>
    distinct(genspe) |>
    anti_join(species |> distinct(genspe)) |>
    summarize(n = n()) |>
    pull(n) == 0
)
```

# Quik Viz

## Stems by Species

```{r}
trees |>
  filter(gesp != "SNAG") |>
  group_by(gesp) |>
  summarize(n = n()) |>
  slice_max(n, n = 10) |>
  ggplot(aes(fct_reorder(gesp, desc(n)), n)) +
  geom_col() +
  theme_bw() +
  ggtitle("Luna's Laser Plots: Number of Trees by Species") +
  xlab("Genus / Species") +
  ylab("Number of Stems")
```

## Stems by Subplot

```{r}
trees |>
  filter(gesp != "SNAG") |>
  mutate(distance_from_edge = case_when(
    subplot == 1 ~ "0m - 10m",
    subplot == 2 ~ "10m - 20m",
    subplot == 3 ~ "20m - 30m"
  )) |>
  group_by(plot_id, distance_from_edge) |>
  summarize(n = n(), .groups = "drop") |>
  ggplot(aes(distance_from_edge, n)) +
  geom_boxplot(notch = TRUE) + 
  theme_bw() +
  ggtitle("Luna's Laser Plots: stems by subplot") +
  ylab("Number of Stems >= 10cm DBH") +
  xlab("Distance from Edge")
```

# DBH by Subplot

```{r}
trees |>
  filter(gesp != "SNAG") |>
  mutate(distance_from_edge = case_when(
    subplot == 1 ~ "0m - 10m",
    subplot == 2 ~ "10m - 20m",
    subplot == 3 ~ "20m - 30m"
  )) |>
  ggplot(aes(distance_from_edge, dbh)) +
  geom_boxplot(notch = TRUE) + 
  theme_bw() +
  ggtitle("Luna's Laser Plots: DBH by subplot") +
  ylab("Diameter (cm)") +
  xlab("Distance from Edge")
```

# Biomass

## Formulae

Harvard Forest code is from Danelle

Allodb can fill in the few species for which Harvard Forest doesn't
have biomass, e.g. Acer platanoides and Quercus macrocarpa.

biomass_AGBMg moved to R/functions.R

spCoef is:
```{r}
merged_genJen_HF3_20260611 <- read_csv(file.path(
  "data/raw",
  "merged_genJen_HF3_20260611.csv"
)) |>
  # There are duplicate rows; pick the first one
  group_by(genspe) |>
  slice_head(n = 1) |>
  ungroup()
```

## Biomass

Let's try it

```{r}
biomass <- trees |>
  biomass_AGBMg(merged_genJen_HF3_20260611)
```

## Missing Values

Did anything other than SNAGs not get biomass?

```{r}
biomass |>
  filter(gesp != "SNAG") |>
  filter(is.na(AGBMg)) |>
  distinct(gesp)
```

## Outliers

who is the outlier?

```{r}
biomass |>
  group_by(plot_id, subplot) |>
  summarize(AGBMg = sum(AGBMg, na.rm = TRUE), .groups = "drop") |>
  mutate(AGBMg_ha = AGBMg * 10000 / 100) |> # hectare is 10000m2, we have 100m2
  slice_max(AGBMg, n = 1)
```

## Viz

```{r}
biomass |>
  group_by(plot_id, subplot) |>
  summarize(AGBMg = sum(AGBMg, na.rm = TRUE), .groups = "drop") |>
  mutate(
    subplot = ordered(subplot),
    AGBMg_ha = AGBMg * 10000 / 100
  ) |>
  left_join(plots, by = join_by(plot_id)) |>
  ggplot(aes(subplot, AGBMg_ha)) +
  geom_boxplot(outlier.size = 0.25, notch = TRUE) +
  facet_wrap(c(~aspect, ~adjacent), labeller = label_both) +
  theme_bw() +
  ggtitle("Subplot Biomass by Aspect and Adjacency") +
  ylab(bquote("Biomass" ~ (Mg %.% ha^-1))) +
  xlab("Subplot")
```


## Biomass by Species

```{r}
biomass |>
  filter(gesp != "SNAG") |>
  group_by(gesp) |>
  summarize(AGBMg = sum(AGBMg)) |>
  slice_max(AGBMg, n = 10) |>
  ggplot(aes(fct_reorder(gesp, desc(AGBMg)), AGBMg)) +
  geom_col() +
  theme_bw() +
  ggtitle("Luna's Laser Plots: Biomass by Species") +
  xlab("Genus / Species") +
  ylab("Biomass (Mg)")
```

## Biomass Comparison

How much does it matter whether we get the species right?

### Red vs. Sugar Maple

Both of these use formula form 3, which uses b1 and b2.

```{r}
biomass_function_for_species <- function(genspe) {
  row <- merged_genJen_HF3_20260611 |>
    filter(genspe == !!genspe)
  b1 <- row |> pull(b1)
  b2 <- row |> pull(b2)
  b3 <- row |> pull(b3)
  b4 <- row |> pull(b4)
  
  formulaType <- row |> pull(formulaType)
  
  formula <- switch(
    as.character(formulaType),
    "1" = function(dbh) b1 * (dbh^b2),
    "3" = function(dbh) exp(b1 + b2 * log(dbh)),
    "5" = function(dbh) (b1 * (dbh^b2)) / 1000,
    "7" = exp((b1 + b2 * log(dbh)) / 1000),
  )

  return(formula)
}

red_maple <- biomass_function_for_species("ACERUB")
sugar_maple <- biomass_function_for_species("ACESAC")

ggplot() +
  xlim(10, 50) +
  geom_function(aes(color = "Red Maple"), fun = red_maple) +
  geom_function(aes(color = "Sugar Maple"), fun = sugar_maple) +
  theme_bw() +
  scale_color_manual(
    name = "Species",
    values = c("Red Maple" = "red", "Sugar Maple" = "blue")
  ) +
  ggtitle("Biomass by DBH") +
  xlab("DBH (cm)") +
  ylab("Biomass (Kg)")

```

### Red Oak vs. Basswood

```{r}
red_oak <- biomass_function_for_species("QUERUB")
basswood <- biomass_function_for_species("TILAME")

ggplot() +
  xlim(10, 50) +
  geom_function(aes(color = "Red Oak"), fun = red_oak) +
  geom_function(aes(color = "Basswood"), fun = basswood) +
  theme_bw() +
  scale_color_manual(
    name = "Species",
    values = c("Red Oak" = "red", "Basswood" = "blue")
  ) +
  xlab("DBH (cm)") +
  ylab("Biomass (Kg)")

```

### Black Birch vs. Black Cherry

```{r}
black_birch <- biomass_function_for_species("BETLEN")
black_cherry <- biomass_function_for_species("PRUSER")

ggplot() +
  xlim(10, 50) +
  geom_function(aes(color = "Black Birch"), fun = black_birch) +
  geom_function(aes(color = "Black Cherry"), fun = black_cherry) +
  theme_bw() +
  scale_color_manual(
    name = "Species",
    values = c("Black Birch" = "red", "Black Cherry" = "blue")
  ) +
  xlab("DBH (cm)") +
  ylab("Biomass (Kg)")

```


# Plots

## Edge Angle

We're not consistent about whether we measure east-to-west or west-to-east.
Fix it: GIS prefers that exterior use counterclockwise winding direction.

So, if a plot is south facing, the edge edge should go west-to-east (<180);
if a plot is north facing, the edge edge should go east-to-west (>180).

```{r}
magnetic_declination <- -13.5464

# correct for magnetic declination
plots$edge_angle_true <- (plots$edge_angle + magnetic_declination) %% 360

# check and correct plot edge bearing
plots <- plots |>
  mutate(
    long_axis_bearing = (edge_angle_true + 90) %% 360,
    north_component   = cos(long_axis_bearing * pi / 180),
    needs_flip = case_when(
      aspect == "N" & north_component >= 0 ~ TRUE,   # would extend north instead of south
      aspect == "S" & north_component <= 0 ~ TRUE,   # would extend south instead of north
      TRUE ~ FALSE
    ),
    edge_angle_fixed = if_else(needs_flip,
                              (edge_angle_true + 180) %% 360,
                              edge_angle_true) |> round()
  )

```

```{r}
plots <- plots |>
  mutate(edge_angle_fixed = case_when(
    aspect == "N" & edge_angle >= 180 ~ edge_angle,
    aspect == "N" & edge_angle < 180 ~ edge_angle + 180,
    aspect == "S" & edge_angle >= 180 ~ edge_angle - 180,
    aspect == "S" & edge_angle < 180 ~ edge_angle
  ))
```


