---
title: "05_ForestGEO"
output: html_document
---

# Setup

```{r setup}
library(tidyverse)
library(terra)
library(tidyterra)
library(sf) # for polygon manipulation
library(allodb)

for (script in list.files(path = "R", pattern = "*.[Rr]", full.names = TRUE)) {
  source(script)
}
```

# Datums

## Species

```{r}
species <- readxl::read_xlsx("data/raw/Edge Plot Data Sheet.xlsx", sheet = "Tree Species") |>
  mutate(genspe= paste0(toupper(substr(GENUS, 1, 3)), toupper(substr(SPECIES, 1, 3))))
```

## Laser Plots

```{r}
# Laser Plot Data
lp_plots <- readxl::read_xlsx("data/raw/Edge Plot Data Sheet.xlsx", sheet = "Plot Data")
lp_trees <- readxl::read_xlsx("data/raw/Edge Plot Data Sheet.xlsx", sheet = "Tree Data") |>
  left_join(species, by = join_by(gesp))
```

## Harvard Forest

```{r}
# HF point-of-interest
harvard_forest <- terra::vect(file.path(
  "/vsizip",
  "data/processed",
  "harvard_forest.shp.zip",
  "harvard_forest.shp.shp"
)) |>
  as.data.frame(geom = "XY") |>
  select(lon = x, lat = y) |>
  as.list()
```

## ForestGEO


### Plot Layout

ForestGEO plot layout starts in the southwest corner. The first
two digits of the quadrat ID are the column, starting with "01"
in the west and ending with "35" in the east. The second two
digits of the quadrat are the row, starting with "01" in the south
and ending with "25" in the north. The first digit of the subquad
is the column, starting with "1" in the west and ending with "2"
in the east. The second digit of the subquad is the row, starting
with "1" in the south and ending with "2" in the north.

So, the most southwest subquad is in quad 01 01, subquad 1 1,
creating a complete id of 010111, often styled 0101_11. The
northwesternmost subquad is 01 25 1 2, or 012512. The northeasternmost
subquad is 352522, and the southeasternmost is 350121.

Here's an attempt to lay it out:

```{r forestgeo_plot_layout}
fg_layout <- tribble(
  ~a, ~b, ~c, ~d, ~e, ~f, ~g, ~g,
  "xx25", "12", "22", "12", "22", "...", "12", "22",
  "xx25", "11", "21", "11", "21", "...", "11", "21",
  "...", "...", "...", "...", "...", "...", "...", "...",
  "xx02", "12", "22", "12", "22", "...", "12", "22",
  "xx02", "11", "21", "11", "21", "...", "11", "21",
  "xx01", "12", "22", "12", "22", "...", "12", "22",
  "xx01", "11", "21", "11", "21", "...", "11", "21",
  "", "01yy", "01yy", "02yy", "02yy", "...", "35yy", "35yy"
)
knitr::kable(fg_layout, col.names = NULL)
```

So we can get to subquad row and column as:

* row = quad row * 2 + subquad row - 1
* col = quad col * 2 + subquad col - 1

```{r}
fg_row_col <- expand_grid(
    quad_col = 1:35,
    quad_row = 1:25,
    subq_col = 1:2,
    subq_row = 1:2
  ) |>
  mutate(
    row = (quad_row - 1) * 2 + subq_row,
    col = (quad_col - 1) * 2 + subq_col,
    quadrat = sprintf("%02d%02d", quad_col, quad_row),
    subquad = sprintf("%d%d", subq_col, subq_row)
  ) |>
  select(-quad_col, -quad_row, -subq_col, -subq_row)
```

We also want the layout of quadrat and subquad in UTM 18N.

```{r}
fg_plots <- terra::vect(file.path(
  "data/raw/ForestGEO",
  "quadsSubquadsOutFixed_UTM.shp"
))
```

### Trees

```{r}
fg_stem_to_subquad <- read_csv(
  "data/raw/ForestGEO/stemTagQuadSubquad_cw.csv",
  col_types = cols(
    quadSubquad = col_character(),
    quadrat = col_character(),
    subquad = col_character(),
    tag = col_character(),
    StemTag = col_character()
  )
) |>
  # Add back leading zeroes that get removed if strings that look like
  # numbers get processed as numbers.
  mutate(
    quadrat = sprintf("%04d", as.integer(quadrat)),
    tag = sprintf("%06d", as.integer(tag)),
    StemTag = sprintf("%06d", as.integer(StemTag)),
    quadSubquad = paste0(quadrat, "_", subquad)
  ) |>
  # There are a lot of duplicates!
  # Take the firest one.
  group_by(tag, StemTag) |>
  slice_head(n = 1) |>
  ungroup()

fg_species <- read_csv(
  "data/raw/ForestGEO/hf253-02-species-codes.csv"
) |>
  mutate(
    genspe = paste0(
      toupper(substr(genus, 1, 3)),
      toupper(substr(species, 1, 3))
    )
  )

fg_trees <- read_csv(
  "data/raw/ForestGEO/hf_census_all_v4_2026-03-06.csv"
) |>
  left_join(
    fg_stem_to_subquad |>
      group_by(StemTag) |>
      arrange(tag) |>
      slice_head(n = 1) |>
      ungroup() |>
      select(StemTag, quadrat_stem = quadrat, subquad_stem = subquad),
    by = join_by(StemTag)
  ) |>
  left_join(
    fg_stem_to_subquad |>
      group_by(tag) |>
      slice_head(n = 1) |>
      ungroup() |>
      select(tag, quadrat_tag = quadrat, subquad_tag = subquad),
    by = join_by(tag)
  ) |>
  mutate(
    quadrat = coalesce(quadrat, quadrat_stem, quadrat_tag),
    subquad = coalesce(subquad_stem, subquad_tag)
  ) |>
  select(-quadrat_stem, -quadrat_tag, -subquad_stem, -subquad_tag) |>
  mutate(
    status = if_else(DFstatus == "alive", "A",
      if_else(DFstatus == "stem dead", "D",
        if_else(str_detect(codes, "D"), "D",
          NA
        )
      )
    )
  ) |>
  left_join(
    fg_species,
    by = join_by(sp)
  ) |>
  left_join(
    fg_row_col,
    by = join_by(quadrat, subquad)
  )
```

### Biomass

We need the species-specific allometric equations:
```{r}
merged_genJen_HF3_20260611 <- read_csv(
  file.path("data/raw", "merged_genJen_HF3_20260611.csv"),
  col_types = cols(
    b3 = col_number(),
    b4 = col_number()
  )
) |>
  # There are duplicate rows; pick the first one
  group_by(genspe) |>
  slice_head(n = 1) |>
  ungroup()
```

The biomass function is loaded in #setup

```{r}
fg_biomass <- fg_trees |>
  rename(GENUS = genus, SPECIES = species) |>
  biomass_AGBMg(merged_genJen_HF3_20260611)

```

Aggregate biomass by subquad

```{r}
fg_biomass_by_subquad <- fg_biomass |>
  filter(
    CensusID == 3, # Only the 2025 census
    status == "A", # Only live trees
    dbh >= 10 # Only trees >= 10cm dbh
  ) |>
  group_by(row, col) |>
  summarize(
    AGBMg = sum(AGBMg, na.rm = TRUE),
    .groups = "drop"
  )
```

Is it sane?

```{r}
fg_biomass_by_subquad |>
  select(row, col, AGBMg) |>
#  full_join(fg_row_col |> select(row, col)) |>
  ggplot(aes(col, row, fill = AGBMg)) +
  geom_raster(hjust = 0, vjust = 0) +
  scale_fill_viridis_c(na.value = "black") +
  theme_bw() +
  ggtitle("Harvard Forest ForestGEO Biomass per 10m x 10m subquad")
```

```{r}
fg_biomass_by_subquad |>
  select(row, col, AGBMg) |>
  ggplot(aes(AGBMg)) +
  geom_histogram(binwidth = 1) +
  theme_bw() +
  ggtitle("ForestGEO Biomass Distribution") +
  xlab("Above Ground Biomass (Mg)") +
  ylab("Count")
```

### Canopy Height

NEON has a canopy height product. It's 1m x 1m height in fractional
meters, CRS = EPSG:32618 = WGS 84 / UTM Zone 18N.

Terra is not particularly smart about managing virtual rasters, so
let's figure out which tiles overlap with ForestGEO. So, find the
min/max extent of the ForestGEO plot.

```{r}
fg_ext <- fg_plots |>
  terra::ext()
# This is in meters, and NEON uses 1km x 1km pixels aligned to UTM18N,
# so we can figure out exactly which tiles we need
fg_ext_xmin <- floor(fg_ext$xmin / 1000) * 1000
fg_ext_xmax <- (floor(fg_ext$xmax / 1000) + 1) * 1000
fg_ext_ymin <- floor(fg_ext$ymin / 1000) * 1000
fg_ext_ymax <- (floor(fg_ext$ymax / 1000) + 1) * 1000
```

```{r}
fg_chm <- file.path(
    "data",
    "raw",
    "NEON_struct-ecosystem",
    "NEON.D01.HARV.DP3.30015.001.2024-08.basic.20250515T151700Z.PROVISIONAL"
  ) |>
  list.files(
    pattern = "\\.tif$",
    full.names = TRUE
  ) |>
  tibble(path = _) |>
  mutate(
    fname = basename(path),
    easting = as.numeric(str_extract(fname, "(?<=_)\\d{6}(?=_\\d{7}_)")),
    northing = as.numeric(str_extract(fname, "(?<=_\\d{6}_)\\d{7}(?=_)"))
  ) |>
  filter(
    easting >= fg_ext_xmin & easting < fg_ext_xmax,
    northing >= fg_ext_ymin & northing < fg_ext_ymax
  ) |>
  pull(path) |>
  terra::vrt()
names(fg_chm) <- "canopy_height"
```

Aggregate canopy height to subquads

```{r}
fg_chm_by_subquad <- fg_chm |>
  terra::extract(fg_plots, fun = mean, na.rm = TRUE, bind = TRUE)
```

Is it sane?

```{r}
fg_chm_by_subquad |>
  as.data.frame() |>
  mutate(subquad = as.character(subquad)) |>
  select(quadrat, subquad, canopy_height) |>
  left_join(fg_row_col, by = join_by(quadrat, subquad)) |>
  select(row, col, canopy_height) |>
  ggplot(aes(col, row, fill = canopy_height)) +
  geom_raster(hjust = 0, vjust = 0) +
  scale_fill_viridis_c(na.value = "black") +
  theme_bw() +
  ggtitle("Harvard Forest ForestGEO Mean Canopy Height per 10m x 10m subquad") +
  labs(fill = "Canopy\nHeight (m)")
```

And a distribution of canopy height, only for pixels with biomass

```{r}
fg_chm_by_subquad |>
  as.data.frame() |>
  select(quadrat, subquad, canopy_height) |>
  mutate(subquad = as.character(subquad)) |>
  left_join(
    fg_row_col,
    by = join_by(quadrat, subquad)
  ) |>
  semi_join(
    fg_biomass_by_subquad |>
      select(row, col),
    by = join_by(row, col)
  ) |>
  ggplot(aes(col, row, fill = canopy_height)) +
  geom_raster(hjust = 0, vjust = 0) +
  scale_fill_viridis_c(na.value = "black") +
  theme_bw() +
  ggtitle("Harvard Forest ForestGEO Mean Canopy Height per 10m x 10m subquad") +
  labs(fill = "Canopy\nHeight (m)")

```

```{r}
fg_chm_by_subquad |>
  as.data.frame() |>
  select(quadrat, subquad, canopy_height) |>
  mutate(subquad = as.character(subquad)) |>
  left_join(
    fg_row_col,
    by = join_by(quadrat, subquad)
  ) |>
  semi_join(
    fg_biomass_by_subquad |>
      select(row, col),
    by = join_by(row, col)
  ) |>
  ggplot(aes(canopy_height)) +
  geom_histogram(binwidth = 1) +
  theme_bw() +
  ggtitle("ForestGEO Canopy Height Distribution") +
  xlab("Canopy Height (m)") +
  ylab("Count")
```

## Edges

### Biomass

```{r}
lp_biomass <- lp_trees |>
  biomass_AGBMg(merged_genJen_HF3_20260611)
```

```{r}
lp_biomass_by_subplot <- lp_biomass |>
  filter(gesp != "SNAG") |>
  group_by(plot_id, subplot) |>
  summarize(AGBMg = sum(AGBMg, na.rm = TRUE), .groups = "drop")
```

### Plot Polygons

```{r}
edge_plot_poly_input <- lp_plots |>
  mutate(edge_angle = edge_angle + 14) |>
  select(plot_id, lon, lat, edge_angle) |>
  terra::vect(crs = crs("epsg:4326")) |>
  terra::project(neon_chm) |>
  as.data.frame(geom = "XY")

# Create a simple feature (sf) object with all the subplot polygons
lp_poly_sf <- do.call(rbind, Map(
  edge_plot_poly,
  edge_plot_poly_input$x, edge_plot_poly_input$y,
  edge_plot_poly_input$edge_angle, edge_plot_poly_input$plot_id
))
st_crs(lp_poly_sf) <- st_crs(neon_chm)

# Create a SpatVector from them so we can do things with it
lp_subplots <- lp_poly_sf |>
  terra::vect()
```

Check 'em out
```{r}
lp_poly_sf |>
  head(3) |>
  terra::vect() |>
  ggplot() +
  geom_spatvector() +
  coord_sf()
```

### Canopy Height

```{r}
lp_ext <- lp_subplots |>
  terra::ext()
# This is in meters, and NEON uses 1km x 1km pixels aligned to UTM18N,
# so we can figure out exactly which tiles we need
lp_ext_xmin <- floor(lp_ext$xmin / 1000) * 1000
lp_ext_xmax <- (floor(lp_ext$xmax / 1000) + 1) * 1000
lp_ext_ymin <- floor(lp_ext$ymin / 1000) * 1000
lp_ext_ymax <- (floor(lp_ext$ymax / 1000) + 1) * 1000
```

```{r}
lp_chm <- file.path(
    "data",
    "raw",
    "NEON_struct-ecosystem",
    "NEON.D01.HARV.DP3.30015.001.2024-08.basic.20250515T151700Z.PROVISIONAL"
  ) |>
  list.files(
    pattern = "\\.tif$",
    full.names = TRUE
  ) |>
  tibble(path = _) |>
  mutate(
    fname = basename(path),
    easting = as.numeric(str_extract(fname, "(?<=_)\\d{6}(?=_\\d{7}_)")),
    northing = as.numeric(str_extract(fname, "(?<=_\\d{6}_)\\d{7}(?=_)"))
  ) |>
  filter(
    easting >= lp_ext_xmin & easting < lp_ext_xmax,
    northing >= lp_ext_ymin & northing < lp_ext_ymax
  ) |>
  pull(path) |>
  terra::vrt()
names(lp_chm) <- "canopy_height"
```

Aggregate canopy height to subquads

```{r}
lp_chm_by_subplot <- lp_chm |>
  terra::extract(lp_subplots, fun = mean, na.rm = TRUE, bind = TRUE)
```


Is it sane?

Distribution of canopy height

```{r}
lp_chm_by_subplot |>
  as.data.frame() |>
  select(plot_id, subplot_id, canopy_height) |>
  ggplot(aes(canopy_height)) +
  geom_histogram(binwidth = 1) +
  theme_bw() +
  ggtitle("Edge Plot Mean Canopy Height per 10m x 10m subplot")

```

# Compare ForestGEO and Edge Plot

## Canopy Height

```{r}
fg_chm_by_subquad |>
  as.data.frame() |>
  mutate(series = "ForestGEO") |>
  select(series, canopy_height) |>
  union_all(
    lp_chm_by_subplot |>
      as.data.frame() |>
      mutate(series = "Edge") |>
      select(series, canopy_height)
  ) |>
  ggplot(aes(canopy_height, y = after_stat(density), fill = series)) +
  geom_histogram(binwidth = 1, position = "dodge") +
  theme_bw() +
#  scale_fill_viridis_d() +
  ggtitle("Canopy Height Comparison") +
  ylab("Density") +
  xlab("Canopy Height (m)") +
  labs(fill = "Plots")
```
