---
title: "13_Wisconsin"
output: html_document
---

# Setup

```{r setup}
library(tidyverse)
library(terra)
library(tidyterra)
library(sf)
library(sfheaders)
library(ggspatial)
library(patchwork)
library(gdalUtilities) # ogr2ogr
library(maptiles) # for basemap

for (script in list.files(path = "R", pattern = "*.[Rr]", full.names = TRUE)) {
  source(script)
}
```

# Datums

Wisconsin is also home to:

* UNDE - University of Notre Dame Environmental Research Center
* LIRO - Little Rock Lake
* STEI - Steigerwaldt-Chequamegon, but D05_CHEQ_R2_P1_v2
  (in the Chequamegon-Nicolet National Forest, Price County)

These have fewer edges than HARV.

Wisconsin GIS portal: https://data-wi-dnr.opendata.arcgis.com/

Only department of natural resources; does not have real estate records

Real estate records are searchable by an 3rd party service for $9 per search

## State and County Outlines

```{r state_outlines}
wi_outline <- map_data("state", region = "Wisconsin") |>
  sfheaders::sf_polygon(
    obj = _,
    x = "long",
    y = "lat",
    polygon_id = "group"
  ) |>
  `st_crs<-`(sf::st_crs("epsg:4326")) |>
  terra::vect()

wi_counties <- terra::vect(file.path(
   "/vsizip",
   "data/raw/WI",
   "County_Boundaries_24K.zip",
   "County_Boundaries_24K.shp"
))
```


## NEON STEI

STEI = Steigerwaldt-Chequamegon, D05_STEI_TREE_R1_P1_v2

Lincoln County

```{r neon_aop_stei}
neon_aop_stei <- terra::vect(file.path(
   "/vsizip",
   "data/raw",
   "NEON_AOP_flightBoxes_0.zip",
   "AOP_flightBoxes",
   "AOP_flightboxesAllSites.shp"
)) |>
  filter(flightbxID == "D05_STEI_TREE_R1_P1_v2")

neon_aop_stei_buffered <- neon_aop_stei |>
  terra::buffer(1000)
```

## NEON TREE

TREE = Treehaven
Oneida & Langlade Counties

```{r neon_aop_tree}
neon_aop_tree <- terra::vect(file.path(
   "/vsizip",
   "data/raw",
   "NEON_AOP_flightBoxes_0.zip",
   "AOP_flightBoxes",
   "AOP_flightboxesAllSites.shp"
)) |>
  filter(siteID == "TREE")

neon_aop_tree_buffered <- neon_aop_tree |>
  terra::buffer(1000)
```

## Tax Parcels

This is from https://geodata.wisc.edu/catalog/42c3da8b-66a1-4321-bf80-ea1c5ed37cf9

```{r wi_tax_parcels}
wi_lincoln_tax_parcels <- terra::vect(file.path(
  "/vsizip",
  "data/raw/WI",
  "V1200_Wisconsin_Parcels_LINCOLN_SHP.zip",
  "V1200_Wisconsin_Parcels_LINCOLN.shp"
))

wi_oneida_tax_parcels <- terra::vect(file.path(
  "/vsizip",
  "data/raw/WI",
  "V1200_Wisconsin_Parcels_ONEIDA_SHP.zip",
  "V1200_Wisconsin_Parcels_ONEIDA.shp"
))

wi_langlade_tax_parcels <- terra::vect(file.path(
  "/vsizip",
  "data/raw/WI",
  "V1200_Wisconsin_Parcels_LANGLADE_SHP.zip",
  "V1200_Wisconsin_Parcels_LANGLADE.shp"
))

wi_tax_parcels <-  c(
  wi_lincoln_tax_parcels,
  wi_oneida_tax_parcels,
  wi_langlade_tax_parcels
) |>
  terra::vect() |>
  terra::project(neon_aop_stei) |>
  terra::intersect(
    c(neon_aop_stei_buffered, neon_aop_tree_buffered) |>
      terra::vect()
  )
```

```{r}
ggplot() +
  geom_spatvector(
    data = wi_tax_parcels,
    color = "black",
    fill = "transparent"
  ) +
  geom_spatvector(
    data = neon_aop_stei,
    color = "red",
    fill = "transparent"
  ) +
  geom_spatvector(
    data = neon_aop_tree,
    color = "red",
    fill = "transparent"
  ) +
  geom_spatvector(
    data = wi_counties,
    color = "blue",
    fill = "transparent"
  ) +
  coord_sf(
    crs = crs(wi_outline),
    xlim = c(ext(c(neon_aop_stei, neon_aop_tree))$xmin, ext(c(neon_aop_stei, neon_aop_tree))$xmax),
    ylim = c(ext(c(neon_aop_stei, neon_aop_tree))$ymin, ext(c(neon_aop_stei, neon_aop_tree))$ymax)
  ) +
  theme_bw() +
  ggtitle("NEON AOPs in WI") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  annotation_scale(
    location = "br",
    pad_x = unit(0.45, "cm"),
    pad_y = unit(0.1, "cm")
  ) +
  annotation_north_arrow(
    location = "tr",
    width = unit(0.75, "cm"),
    pad_x = unit(-0.15, "cm"),
    pad_y = unit(0.25, "cm"),
    style = north_arrow_minimal
  )

```

```{r}
wi_tax_parcels |>
  terra::writeVector("data/processed/WI_Tax_parcels.shp.zip")
```

