---
title: "14_Maryland"
output: html_document
---

# Setup

```{r setup}
library(tidyverse)
library(terra)
library(tidyterra)
library(sf)
library(sfheaders)
library(ggspatial)
library(patchwork)
library(gdalUtilities) # ogr2ogr
library(maptiles) # for basemap

for (script in list.files(path = "R", pattern = "*.[Rr]", full.names = TRUE)) {
  source(script)
}
```

# Datums

## NEON SERC

SERC - Smithsonian Environmental Research Center

```{r neon_aop_scbi}
neon_aop_serc <- terra::vect(file.path(
   "/vsizip",
   "data/raw",
   "NEON_AOP_flightBoxes_0.zip",
   "AOP_flightBoxes",
   "AOP_flightboxesAllSites.shp"
)) |>
  filter(siteID == "SERC")

neon_aop_serc_buffered <- neon_aop_serc |>
  terra::buffer(1000)
```


## Tax Parcels

Tax data at https://maryland-department-of-planning-data-hub-maryland.hub.arcgis.com/#gis-downloads

Downloads are by county.

SERC is in Anne Arundel county.

```{r md_tax_parcels}
md_tax_parcels <- terra::vect(file.path(
  "/vsizip",
  "data/raw/MD",
  "ANNEparcels0526.zip",
  "ANNEPOLY.shp"
)) |>
  terra::project(neon_aop_serc) |>
  terra::intersect(neon_aop_serc_buffered)
```

Because Annapolis is in the NEON AOP, there are a bazillion tiny
tax parcels. Create a ha attribute so this can be filtered by size.

```{r md_tax_parcels_1ha}
md_tax_parcels$ha <- md_tax_parcels |>
  terra::expanse(unit = "ha")
```

# Preview


## State and County Outlines

```{r state_outlines}
md_outline <- map_data("state", region = "Maryland") |>
  sfheaders::sf_polygon(
    obj = _,
    x = "long",
    y = "lat",
    polygon_id = "group"
  ) |>
  `st_crs<-`(sf::st_crs("epsg:4326")) |>
  terra::vect()

md_counties <- map_data("county", region = "Maryland") |>
  sfheaders::sf_polygon(
    obj = _,
    x = "long",
    y = "lat",
    polygon_id = "group"
  ) |>
  `st_crs<-`(sf::st_crs("epsg:4326")) |>
  terra::vect()
```


```{r}
ggplot() +
  geom_spatvector(
    data = md_counties,
    color = "blue",
    fill = "transparent"
  ) +
  geom_spatvector(
    data = md_tax_parcels,
    color = "black",
    fill = "transparent"
  ) +
  geom_spatvector(
    data = neon_aop_serc,
    color = "red",
    fill = "transparent"
  ) +
  coord_sf(
    crs = crs(md_outline),
    xlim = c(ext(neon_aop_serc)$xmin, ext(neon_aop_serc)$xmax),
    ylim = c(ext(neon_aop_serc)$ymin, ext(neon_aop_serc)$ymax)
  ) +
  theme_bw() +
  ggtitle("NEON AOPs in MD") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  annotation_scale(
    location = "br",
    pad_x = unit(0.45, "cm"),
    pad_y = unit(0.1, "cm")
  ) +
  annotation_north_arrow(
    location = "tr",
    width = unit(0.75, "cm"),
    pad_x = unit(-0.15, "cm"),
    pad_y = unit(0.25, "cm"),
    style = north_arrow_minimal
  )

```


# Write

Write out a small file of tax parcels for use in QGIS

```{r write md_tax_parcels}
md_tax_parcels |>
  terra::writeVector(file.path(
    "data/processed",
    "MD_Tax_parcels.shp.zip"
  ))
```

# Notes

